# knitr document van Steensel lab

# Escaper/repressed TRIP data pre-processing
## Christ Leemans, 14-09-2017

## Introduction
In previous analysis we used SuRE and GROcap to classify promoters in lamina
associated domains as repressed, inactive or escaper. Repressed being actively
repressed, inactive promoters showing no activity even when removed from the lamina
and escapers being equally active in both SuRE and GROcap.

To validate that the escaping behavior is encoded in the promoter sequence,
Marloes created TRIP pools for 4 escaper promoters and 4 repressed promoters.
If escaping behavior is encoded in the sequence, any integration of an escaper
promoter in LADs should be able to escape repression, while integrations of
repressed promoters should be repressed.

Recently the first data came back from the sequencing facility. This contains
3 promoters (2 repressed, 1 escaper), each two biological replicates and two
technical replicates.



```{r}
library(ggplot2)

pseudo_log10 <- function(val_vec){
    Pseud=min(val_vec[val_vec > 0], na.rm=TRUE)/2
    val_vec = val_vec + Pseud
    return(log10(val_vec))
}


file_list = list.files('../cl20170913_TRIP_K562_evsr/', pattern='normalized',
                       full.names=T)
names(file_list) = gsub('.*cDNA.(.*).normalized', '\\1', file_list)

table_list = lapply(file_list, read.table, header=T, stringsAsFactors=F,
                    row.names=1)

info = data.frame(name=names(file_list),
                  promoter=gsub('([0-9]+).*', '\\1', names(file_list)),
                  bio_rep = gsub('[0-9]+([A-B]).*', '\\1', names(file_list)),
                  tech_rep = gsub('[0-9]+[A-B]_rep(.*)', '\\1', names(file_list)),
                  stringsAsFactors=F)
```


```{r, fig.width=10, fig.height=10}
check_reproducability <- function(rep1, rep2, label){
    bc_vec1 = rownames(rep1)
    bc_vec2 = rownames(rep2)
    bc_vec = bc_vec1[bc_vec1 %in% bc_vec2]
    uniq_rep1 = bc_vec1[!bc_vec1 %in% bc_vec2]
    uniq_rep2 = bc_vec2[!bc_vec2 %in% bc_vec1]
    cat(paste0('shared: ', length(bc_vec), '\n'))
    cat(paste0('rep1 unique: ', length(uniq_rep1), '\n'))
    cat(paste0('rep2 unique: ', length(uniq_rep2), '\n'))
    bc_vec = bc_vec[rep1[bc_vec, 'gDNA_count'] > 100 | rep2[bc_vec, 'gDNA_count'] > 100]
    cat(paste0('above threshold (> 100 gDNA either rep): ', length(bc_vec), '\n'))
    cat(paste0('rep1 unique > 100: ', length(which(rep1[uniq_rep1, 'gDNA_count'] > 100)),
               '\n'))
    cat(paste0('rep2 unique > 100: ', length(which(rep2[uniq_rep2, 'gDNA_count'] > 100)),
               '\n'))
    norm_data = data.frame(rep1 = rep1[bc_vec, 'normalized_by_gDNA'],
                          rep2 = rep2[bc_vec, 'normalized_by_gDNA'],
                          gdna1 = rep1[bc_vec, 'gDNA_cpm'],
                          gdna2 = rep2[bc_vec, 'gDNA_cpm'],
                          row.names=bc_vec)
    rep_cor = cor(norm_data$rep1, norm_data$rep2)
    print(ggplot(norm_data, aes(x=pseudo_log10(rep1), y=pseudo_log10(rep2))) +
         geom_point(alpha=0.4, size=0.4) +
         ggtitle(paste0(label, '; cor=', rep_cor)))

    print(ggplot(norm_data, aes(x=pseudo_log10(gdna1 + gdna2),
                         y=pseudo_log10(rep1)-pseudo_log10(rep2))) +
         geom_point(alpha=0.3, size=0.3) +
         ggtitle(paste0(label,
                        ' replicate difference vs total gDNA count; cor=',
                        rep_cor)))
    return(bc_vec)
}

add_density <- function(rep1, rep2, bc_vec, label){
    combined_data = data.frame(cDNA = rep1[bc_vec, 'cDNA_count'] +
                                      rep2[bc_vec, 'cDNA_count'],
                               gDNA = rep1[bc_vec, 'gDNA_count'] +
                                      rep2[bc_vec, 'gDNA_count'],
                               color=label)
    combined_data$normalized = (combined_data$cDNA / sum(combined_data$cDNA)) /
                               (combined_data$gDNA / sum(combined_data$gDNA))
    return(geom_density(data = combined_data, adjust=1/2))
}


density_plot = ggplot(data.frame(), aes(x=pseudo_log10(normalized), color=color))
bc_list = list()
# for (prom in unique(info$promoter)){
class_vec = c('8'='escaper', '11'='repressed')
COL_vec = c(repressed="#A020F0", escaper="#FFA500")
for (prom in c('8', '11')){
    for (bio in unique(info$bio_rep)){
        cat(paste0(prom, bio, ':\n'))
        name_vec = info[info$promoter==prom & info$bio==bio, 'name']
        rep1 = table_list[[name_vec[1]]]
        rep2 = table_list[[name_vec[2]]]
        class = as.character(class_vec[prom])
        label = paste0(prom, bio)
        bc_vec = check_reproducability(rep1, rep2, label)
        bc_list[[label]] = bc_vec
        density_plot = density_plot + add_density(rep1, rep2, bc_vec, class)

    }
}
density_plot + scale_color_manual(values=c(COL_vec, COL_vec))
```

**conclusions:**

For promoter 11 and 8 the reproducibility between technical replicates looks
reasonable, unfortunately promoter 6 is not as reproducible. There seems to be a group
of promoters that show higher variation between replicates, would be interesting
to know what defines this group.

The overall density plot of barcode expression levels is really similar between
biological replicates. This is really nice.

The mayor difference in distribution between promoter 11 and promoter 8 is the
subgroup of barcodes in promoter 8 with a slightly lower expression and the
amount of barcodes that are not expressed. Promoter 8 is an escaper promoter, while
promoter 11 is a repressed promoter. This agrees with previous observations,
that although escaper promoters are equally active as iLAD promoters,
initiation/elongation of POL2 is more difficult in escapers.

```{r, fig.width=10, fig.height=10}



density_plot = ggplot(data.frame(), aes(x=pseudo_log10(normalized), color=color))
for (prom in c('11', '8')){
  for (bio in unique(info$bio_rep)){
      cat(paste0(prom, bio, ':\n'))
      name_vec = info[info$promoter==prom & info$bio==bio, 'name']
      rep1 = table_list[[name_vec[1]]]
      rep2 = table_list[[name_vec[2]]]
      label = paste0(prom, bio)
      bc_vec = bc_list[[label]]
      density_plot = density_plot + add_density(rep1, rep2, bc_vec, prom)
  }
}
file_list = list.files('../cl20160914_TTRIP_K562/', pattern='KRAB_GAL4.*_[1-2]+_[1-2].normalized',
                       full.names=T)
names(file_list) = gsub('.*cDNA.(.*).normalized', '\\1', file_list)

table_list = lapply(file_list, read.table, header=T, stringsAsFactors=F,
                    row.names=1)

info = data.frame(name=names(file_list),
                  promoter=gsub('(.*)_.*_.*', '\\1', names(file_list)),
                  day = gsub('.*_(.*)_.*', '\\1', names(file_list)),
                  tech_rep = gsub('.*_.*_(.*)', '\\1', names(file_list)),
                  stringsAsFactors=F)

for (prom in unique(info$promoter)){
  for (day in unique(info$day)){
      cat(paste0(prom, day, ':\n'))
      name_vec = info[info$promoter==prom & info$day==day, 'name']
      rep1 = table_list[[name_vec[1]]]
      rep2 = table_list[[name_vec[2]]]
      label = paste0(prom,'_day', day)
      bc_vec = check_reproducability(rep1, rep2, label)
      bc_list[[label]] = bc_vec
      density_plot = density_plot + add_density(rep1, rep2, bc_vec, label)

  }
}
density_plot

```

**conclusions:**
Compared to KRAB tethering experiment, the escaper promoters have similar amount
of barcodes with 0 expression (mind the pseudo count). But the expression range
of the pGK promoter in control conditions is more narrow. Day 2 with GAL4-KRAB
condition on the other hand shows 0 expression barcodes similar to repressed promoters
