# knitr document van Steensel lab

# Gene expression correlation between transcription factors and potential target genes
## Christ Leemans, 06-03-2017

## Introduction
In our SURE vs endogenous expression analysis in LADs we identified 3 groups of promoters (escaper, inactive and repressed). We are now looking for a transcription factor that is able to instigate the escaper mechanism in which a promoter locally detaches from the lamina and shows activity. One line of evidence would be that the particular transcription factor correlates with the expression of our escaping promoters. We can use that to filter out TF's of interest from the JASPAR database. So the question for this report is, can we create a matrix of correlation between the expression of our JASPAR TFs and potential target genes in the fantom5 database.




## biomaRt on cluster is messed up, so I do couple jaspar to entrez-gene id first and save it

```r
tf_table = read.table('../raw_data/tf_table.txt', sep='\t', row.names=1, stringsAsFactors=F)
colnames(tf_table) = c('name', 'species', 'class', 'family')

tf_translation = ddply(tf_table[,c('species','name')], .(name),
                       function(x){
                         symbol = gsub('[(]var.[0-9][)]','', x[,2])
                         cbind(x,symbol=unlist(strsplit(symbol, '::')))
                       })


humanMart = useMart(biomart = 'ensembl', dataset = 'hsapiens_gene_ensembl')
bm_eg = getBM(attributes=c('entrezgene','hgnc_symbol'),
              filters='hgnc_symbol', values=tf_translation$symbol, mart=humanMart)
tf_translation$gene_id = bm_eg[match(tf_translation$symbol, bm_eg$hgnc_symbol), 'entrezgene']

species = list(c('Mus musculus', 'mmusculus', 'mgi_symbol'),
               c('Rattus norvegicus', 'rnorvegicus', ''),
               c('Oryctolagus cuniculus', 'ocuniculus'), 
               c('Meleagris gallopavo', 'mgallopavo'), c('Gallus gallus', 'ggallus'))

translate <- function(species, dataset, symbol_name, translation_table, humanMart){
  mart = useMart(biomart='ensembl', dataset=dataset)
  symbol_vec = translation_table$symbol[translation_table$species==species]
  bm_eg = getBM(attributes=c('ensembl_gene_id', symbol_name),
                filters=symbol_name, values=symbol_vec, mart=mart)
  bm_h = getBM(attributes=c('ensembl_gene_id', 'hsapiens_homolog_ensembl_gene',
                             'hsapiens_homolog_orthology_type', 'hsapiens_homolog_subtype',
                             'hsapiens_homolog_perc_id'),
                filters='ensembl_gene_id', values=bm_eg$ensembl_gene_id, mart=mart)
  bm_h = bm_h[bm_h$hsapiens_homolog_ensembl_gene!='', ]
  bm = ddply(bm_h, .(ensembl_gene_id), function(x, bm_eg){
    i = which(x$hsapiens_homolog_perc_id==max(x$hsapiens_homolog_perc_id))
    result = bm_eg[which(bm_eg$ensembl_gene_id==x$ensembl_gene_id[1]), ]
    result = c(result, hsapiens_gene_id=x[i,'hsapiens_homolog_ensembl_gene'])
    return(unlist(result))
  }, bm_eg=bm_eg)
  bm_hg = getBM(attributes=c('entrezgene','hgnc_symbol', 'ensembl_gene_id'),
                filters='ensembl_gene_id', values=bm$hsapiens_gene_id, mart=humanMart)
  bm_match = match(bm_hg$ensembl_gene_id, bm$hsapiens_gene_id)
  bm = data.frame(entrezgene=bm_hg$entrezgene, symbol=bm[bm_match, symbol_name])
  return(bm)
}

bm = translate('Mus musculus', 'mmusculus_gene_ensembl', 'mgi_symbol', tf_translation, humanMart)
tf_match = match(bm$symbol, tf_translation$symbol)
tf_translation[tf_match[!is.na(tf_match)], 'gene_id'] = bm$entrezgene[!is.na(tf_match)]

bm = translate('Rattus norvegicus', 'rnorvegicus_gene_ensembl', 'rgd_symbol', tf_translation, humanMart)
tf_match = match(bm$symbol, tf_translation$symbol)
tf_translation[tf_match[!is.na(tf_match)], 'gene_id'] = bm$entrezgene[!is.na(tf_match)]

bm = translate('Oryctolagus cuniculus', 'ocuniculus_gene_ensembl', 'hgnc_symbol', tf_translation, humanMart)
tf_match = match(bm$symbol, tf_translation$symbol)
tf_translation[tf_match[!is.na(tf_match)], 'gene_id'] = bm$entrezgene[!is.na(tf_match)]

bm = translate('Meleagris gallopavo', 'mgallopavo_gene_ensembl', 'hgnc_symbol', tf_translation, humanMart)
tf_match = match(bm$symbol, tf_translation$symbol)
tf_translation[tf_match[!is.na(tf_match)], 'gene_id'] = bm$entrezgene[!is.na(tf_match)]

bm = translate('Gallus gallus', 'ggallus_gene_ensembl', 'hgnc_symbol', tf_translation, humanMart)
tf_match = match(bm$symbol, tf_translation$symbol)
tf_translation[tf_match[!is.na(tf_match)], 'gene_id'] = bm$entrezgene[!is.na(tf_match)]

write.table(tf_translation, file='../raw_data/cl20160306_tf_translation_entrez.txt', sep='\t')
```

## libraries
```{r}
library(data.table)
library(multicore)
setDTthreads(10)
```


## jaspar-fantom correlation
```{r}
fantom_counts = read.table('~/data/fantom/hg19.cage_peak_phase1and2combined_counts.osc.txt.gz',
                            header=T, stringsAsFactors=F, row.names=1)
fantom_annot = read.table('~/data/fantom/hg19.cage_peak_phase1and2combined_ann.txt.gz',
                           sep='\t', header=T)
load("../raw_data/gencode.sure.160617.rda")
total_vec = fantom_counts[1,]
gencode_transcript_vec = do.call(rbind, strsplit(gencode.sure.160617$name, split='[.]'))[,1]

gencode_fantom = mclapply(gencode_transcript_vec, function(x, association, counts){
        y = grep(x, association)
        if (length(y) > 0){
            return(colSums(counts[y,]))
        }
    }, association=fantom_annot$association_with_transcript, counts=fantom_counts[-1,], mc.cores=5)
names(gencode_fantom) = gencode_transcript_vec
gencode_count = do.call(rbind, gencode_fantom)
gencode_table = as.data.table(gencode_count)
norm_matrix  = sweep(gencode_table, 2, as.numeric(total_vec), '/')
norm_table = as.data.table(t(norm_matrix))
cor_table = norm_table[,lapply(.SD,function(x,dt){cor(dt,x)}, dt=norm_table)]
colnames(cor_table) = rownames(gencode_count)
save(cor_table, file='cl20170426_transcript_correlations.RData')
cor_df = as.data.frame(cor_table)
dimnames(cor_df) = list(rownames(gencode_count), rownames(gencode_count))
save(cor_df, file='cl20170426_transcript_correlations_df.RData')
write.table(cor_df, 'cl20170426_transcript_correlations.txt', quote=F, sep='\t', row.names=T, col.names=T)


enst_table = fantom_table[!is.na(enst_vec), lapply(.SD, sum, na.rm=T), by=enst_vec[!is.na(enst_vec)]] 

tf_table = read.table('../raw_data/cl20160306_tf_translation_entrez.txt', sep='\t', stringsAsFactors=F)


enst_vec = gsub('.*(ENST[0-9]+).*', '\\1', fantom_annot$association_with_transcript)
enst_vec[-grep('ENST', enst_vec)] = NA

fantom_table = data.table(fantom_counts[-1,])
enst_table = fantom_table[!is.na(enst_vec), lapply(.SD, sum, na.rm=T), by=enst_vec[!is.na(enst_vec)]] 
norm_matrix  = sweep(enst_table[,-1], 2, as.numeric(total_vec), '/')
norm_table = as.data.table(t(norm_matrix))
cor_table = norm_table[,lapply(.SD,function(x,dt){cor(dt,x)}, dt=norm_table)]



selection = !is.na(enst_vec) | !is.na(entrez_vec)
fantom_counts = fantom_counts[selection, ]
entrez_vec = entrez_vec[selection]
enst_vec = enst_vec[selection]



enst_table = data.frame(enst=enst_vec[!is.na(enst_vec)],
                        name=rownames(fantom_counts)[!is.na(enst_vec)],
                        stringsAsFactors=F)

fantom_table = data.table(fantom_counts[-1,])
enst_table = fantom_table[!is.na(enst_vec), lapply(.SD, sum, na.rm=T), by=enst_vec[!is.na(enst_vec)]] 

norm_matrix  = sweep(enst_table[,-1], 2, as.numeric(total_vec), '/')

norm_table = as.data.table(t(norm_matrix))
cor_table = norm_table[,lapply(.SD,function(x,dt){cor(dt,x)}, dt=norm_table)]
rownames(cor_table) = enst_table$enst_vec
colnames(cor_table) = enst_table$enst_vec
save(cor_table, file='cl20170424_transcript_correlations.RData')
cor_df = as.data.frame(cor_table)
dimnames(cor_df) = list(enst_table$enst_table, enst_table$enst_table)
save(cor_df, file='cl20170424_transcript_correlations_df.RData')
write.table(cor_df, 'cl20170424_transcript_correlations.txt', quote=F, sep='\t', row.names=T, col.names=T)



cor_gencode = cor_df[gencode_transcript_vec, gencode_transcript_vec]

transcript_id = do.call(rbind, strsplit(gencode.sure.160617$name, spl

test_table = norm_table[,1:10]
start.time <- Sys.time()
test_cor = test_table[,lapply(.SD, function(x,dt){cor(dt,x)}, dt=norm_table)]
end.time <- Sys.time()
time.taken <- end.time - start.time



enst_matrix = ddply(enst_table, .(enst), function(x, fantom_counts){
    colSums(fantom_counts[x$name, ,drop=F])
    }, fantom_counts=fantom_counts, .parallel=T)




norm_matrix = apply(fantom_counts, 2, function(x){
        x / sum(x)
    })
rm(fantom_counts)
tf_in_fantom = tf_table[which(tf_table$gene_id %in% fantom_counts[,1]), ]


cor_table = ddply(tf_in_fantom, .(name), function(x, norm_matrix, entrez_vec){
        id_vec = x$gene_id
        if (length(id_vec) > 1){
            cor_list = list()
            for (i in 1:length(id_vec)){
                match = which(entrez_vec==id_vec[i])
                if (length(match) > 1){
                    cor_list[[i]] = cor(colSums(norm_matrix[match, ]), norm_matrix, method="spearman")
                } else {
                    cor_list[[i]] = cor(norm_matrix[match, ], norm_matrix, method="spearman")
                }
            }
            p_list = do.call(pmin, cor_list)
            return(p_list[1,])
        }
        else {
            match = which(entrez_vec==id_vec)
            if (length(match) > 1){
                return(cor(colSums(norm_matrix[match, ]), norm_matrix, method="spearman"))
            } else {
                return(cor(norm_matrix[match, ], norm_matrix, method="spearman"))
            }
        }
    }, norm_matrix, entrez_vec=entrez_vec, .parallel=T)
rownames(cor_table) = cor_table[,1]
cor_table = cor_table[,-1]

save(cor_table, entrez_vec, file='cl20170404_CAGE_correlations_jaspar_norm.rda')




id = tf_table[tf_table$name=='SPI1', 'gene_id']
match = which(entrez_vec==id)
cor_SPI1 = cor(rowSums(norm_matrix[,match]), norm_matrix[,-match])
SPI1_table = data.frame(cor=cor_SPI1[1,], entrez_gene=fantom_counts[colnames(cor_SPI1),1])
match_vec = match(SPI1_table$entrez_gene, P$entrezgene)
SPI1_table$class = NA
SPI1_table$class[!is.na(match_vec)] = as.character(P[match_vec[!is.na(match_vec)], 'class'])

ggplot(SPI1_table[!is.na(SPI1_table$class), ], aes(x=class, y=cor, color=class)) + 
    geom_violin(alpha=0.3) +
    scale_colour_manual(values=COL_class) +
    geom_point(data=SPI1_table[!is.na(SPI1_table$class)&SPI1_table$class!='iLAD',],
               position=position_jitter(width=0.5), alpha=0.3, size=0.3)


id = tf_table[tf_table$name=='RFX5', 'gene_id']
match = which(entrez_vec==id)
cor_RFX5 = cor(rowSums(count_matrix[,match]), count_matrix[,-match])
RFX5_table = data.frame(cor=cor_RFX5[1,], entrez_gene=fantom_counts[colnames(cor_RFX5),1])
match_vec = match(RFX5_table$entrez_gene, P$entrezgene)
RFX5_table$class = NA
RFX5_table$class[!is.na(match_vec)] = as.character(P[match_vec[!is.na(match_vec)], 'class'])

ggplot(RFX5_table[!is.na(RFX5_table$class), ], aes(x=class, y=cor, color=class)) + 
    geom_violin(alpha=0.3) +
    scale_colour_manual(values=COL_class) +
    geom_point(data=RFX5_table[!is.na(RFX5_table$class)&RFX5_table$class!='iLAD',],
               position=position_jitter(width=0.5), alpha=0.3, size=0.3)


id = tf_table[tf_table$name=='POU2F1', 'gene_id']
match = which(entrez_vec==id)
cor_POU2F1 = cor(rowSums(count_matrix[,match]), count_matrix[,-match])
POU2F1_table = data.frame(cor=cor_POU2F1[1,], entrez_gene=fantom_counts[colnames(cor_POU2F1),1])
match_vec = match(POU2F1_table$entrez_gene, P$entrezgene)
POU2F1_table$class = NA
POU2F1_table$class[!is.na(match_vec)] = as.character(P[match_vec[!is.na(match_vec)], 'class'])

ggplot(POU2F1_table[!is.na(POU2F1_table$class), ], aes(x=class, y=cor, color=class)) + 
    geom_violin(alpha=0.3) +
    scale_colour_manual(values=COL_class) +
    geom_point(data=POU2F1_table[!is.na(POU2F1_table$class)&POU2F1_table$class!='iLAD',],
               position=position_jitter(width=0.5), alpha=0.3, size=0.3)

```


## fantom-fantom correlation
```{r}

cor_table = ddply(enst_table, .(enst), function(x, norm_matrix, enst_table){
        x_vec = colSums(norm_matrix[x$name, ,drop=F])
        cor = ddply(enst_table, .(enst), function(y, norm_matrix, x_vec){
                    y_vec = colSums(norm_matrix[y$name, ,drop=F])
                    cor(x_vec, y_vec, method='pearson')
                }, norm_matrix=norm_matrix, x_vec=x_vec)
        cor_vec = cor[,2]
        names(cor_vec) = cor$enst
        return(cor_vec)
    }, norm_matrix=norm_matrix[!is.na(enst_vec), ], enst_table=enst_table, .parallel=T)



enst_matrix = ddply(enst_table, .(enst), function(x, norm_matrix){
    colSums(norm_matrix[x$name, ,drop=F])
    }, norm_matrix=norm_matrix, .parallel=T)
```
