# knitr document van Steensel lab

# Escaper/repressed TRIP data pre-processing
## Christ Leemans, 14-11-2017

## Introduction
In previous analysis we used SuRE and GROcap to classify promoters in lamina
associated domains as repressed, inactive or escaper. Repressed being actively
repressed, inactive promoters showing no activity even when removed from the lamina
and escapers being equally active in both SuRE and GROcap.

To validate that the escaping behavior is encoded in the promoter sequence,
Marloes created TRIP pools for 4 escaper promoters and 4 repressed promoters.
If escaping behavior is encoded in the sequence, any integration of an escaper
promoter in LADs should be able to escape repression, while integrations of
repressed promoters should be repressed.

Recently the second round of data came back from the sequencing facility. This contains
3 promoters (2 repressed, 1 escaper), each two biological replicates and two
technical replicates.



```{r}
library(ggplot2)
library(rtracklayer)

pseudo_log2 <- function(val_vec){
    Pseud=min(val_vec[val_vec > 0], na.rm=TRUE)/2
    val_vec = val_vec + Pseud
    return(log2(val_vec))
}


file_list = list.files('../cl20171113_TRIP_K562_evsr/cDNA', pattern='normalized',
                       full.names=T)
names(file_list) = gsub('.*cDNA/(.*).normalized', '\\1', file_list)

table_list = lapply(file_list, read.table, header=T, stringsAsFactors=F,
                    row.names=1)

info = data.frame(name=names(file_list),
                  promoter=gsub('([0-9]+).*', '\\1', names(file_list)),
                  bio_rep = gsub('[0-9]+([A-B]).*', '\\1', names(file_list)),
                  tech_rep = gsub('[0-9]+[A-B]_r(.*)', '\\1', names(file_list)),
                  stringsAsFactors=F)
info$promoter = gsub('(.*)_.*_.*', '\\1', info$promoter)
info$bio_rep = gsub('.*_(.*)_.*', '\\1', info$bio_rep)
info$tech_rep = gsub('.*_.*_r(.*)', '\\1', info$tech_rep)
```


```{r, fig.width=10, fig.height=10}
check_reproducability <- function(rep1, rep2, label){
    bc_vec1 = rownames(rep1)
    bc_vec2 = rownames(rep2)
    bc_vec = bc_vec1[bc_vec1 %in% bc_vec2]
    uniq_rep1 = bc_vec1[!bc_vec1 %in% bc_vec2]
    uniq_rep2 = bc_vec2[!bc_vec2 %in% bc_vec1]
    cat(paste0('shared: ', length(bc_vec), '\n'))
    cat(paste0('rep1 unique: ', length(uniq_rep1), '\n'))
    cat(paste0('rep2 unique: ', length(uniq_rep2), '\n'))
    bc_vec = bc_vec[rep1[bc_vec, 'gDNA_count'] > 100 | rep2[bc_vec, 'gDNA_count'] > 100]
    cat(paste0('above threshold (> 100 gDNA either rep): ', length(bc_vec), '\n'))
    cat(paste0('rep1 unique > 100: ', length(which(rep1[uniq_rep1, 'gDNA_count'] > 100)),
               '\n'))
    cat(paste0('rep2 unique > 100: ', length(which(rep2[uniq_rep2, 'gDNA_count'] > 100)),
               '\n'))
    norm_data = data.frame(rep1 = rep1[bc_vec, 'normalized_by_gDNA'],
                          rep2 = rep2[bc_vec, 'normalized_by_gDNA'],
                          gdna1 = rep1[bc_vec, 'gDNA_cpm'],
                          gdna2 = rep2[bc_vec, 'gDNA_cpm'],
                          row.names=bc_vec)
    rep_cor = cor(norm_data$rep1, norm_data$rep2)
    return(bc_vec)
}

add_density <- function(rep1, rep2, bc_vec, label){
    combined_data = data.frame(cDNA = rep1[bc_vec, 'cDNA_count'] +
                                      rep2[bc_vec, 'cDNA_count'],
                               gDNA = rep1[bc_vec, 'gDNA_count'] +
                                      rep2[bc_vec, 'gDNA_count'],
                               color=label)
    combined_data$normalized = (combined_data$cDNA / sum(combined_data$cDNA)) /
                               (combined_data$gDNA / sum(combined_data$gDNA))
    return(geom_density(data = combined_data, adjust=1/2))
}


density_plot = ggplot(data.frame(), aes(x=pseudo_log2(normalized), color=color))
bc_list = list()
# for (prom in unique(info$promoter)){
class_vec = c('6'='repressed', '8'='escaper', '11'='repressed', 'hPGK'='iLAD')
COL_vec = c(repressed="#A020F0", escaper="#FFA500", iLAD="#0077FF")
for (prom in names(class_vec)){
    for (bio in unique(info$bio_rep)){
        cat(paste0(prom, bio, ':\n'))
        name_vec = info[info$promoter==prom & info$bio==bio, 'name']
        rep1 = table_list[[name_vec[1]]]
        rep2 = table_list[[name_vec[2]]]
        class = as.character(class_vec[prom])
        label = paste0(prom, bio)
        bc_vec = check_reproducability(rep1, rep2, label)
        bc_list[[label]] = bc_vec
        density_plot = density_plot + add_density(rep1, rep2, bc_vec, class)
    }
}
density_plot + scale_color_manual(values=c(COL_vec, COL_vec))
```


```{r}


fwd_file_list = list.files('../cl20171113_TRIP_K562_evsr/mapping', pattern='1.table',
                           full.names=T)
names(fwd_file_list) = gsub('.*mapping/(.*).1.table', '\\1', fwd_file_list)


rev_file_list = list.files('../cl20171113_TRIP_K562_evsr/mapping', pattern='2.table',
                           full.names=T)
names(rev_file_list) = gsub('.*mapping/(.*).2.table', '\\1', rev_file_list)


map_info = data.frame(name=names(fwd_file_list),
                      promoter=gsub('([0-9]+).*', '\\1', names(fwd_file_list)),
                      bio_rep = gsub('[0-9]+([A-B]).*', '\\1', names(fwd_file_list)),
                      stringsAsFactors=F)
map_info$promoter = gsub('(.*)_.*', '\\1', map_info$promoter)
map_info$bio_rep = gsub('.*_(.*)', '\\1', map_info$bio_rep)

fwd_table_list = lapply(fwd_file_list, read.table, header=T, stringsAsFactors=F,
                        row.names=1)


rev_table_list = lapply(rev_file_list, read.table, header=T, stringsAsFactors=F,
                        row.names=1)


data_fwd = fwd_table_list[[1]]
data_rev = rev_table_list[[1]]

get_data_table <- function(data_rev, lad_gr, lmnb1_gr){
    map_gr = GRanges(data_rev$seqname, IRanges(data_rev$start_pos,
                                                data_rev$start_pos),
                     strand = data_rev$ori)
    map_gr$LAD = 0
    names(map_gr) = rownames(data_rev)
    o = findOverlaps(map_gr, lad_gr)
    map_gr$LAD[queryHits(o)] = 1
    o = findOverlaps(map_gr, lmnb1_gr)
    map_gr$lamin[queryHits(o)] = lmnb1_gr[subjectHits(o)]$score
    map_gr$ratio1 = data_rev$reads1 / data_rev$total_mapped
    map_gr$ratio2 = data_rev$reads2 / data_rev$total_mapped
    map_gr$avg_mapq = data_rev$mapq_sum1 / data_rev$reads1
    return(map_gr)
}



lad_hmm_gr = import.bed('../../../data/tracks/hg19/cl20161019_LAD_continuous_2state_K562.bed')

lmnb1_10kb = import.bw('/home/t.v.schaik/mydata/proj/3D_nucleus/results/ts170821_GCF4577_K562_p13_LMNB1/results/tracks_norm/LMNB1-10kb.bw')

lad_gr = lad_hmm_gr[lad_hmm_gr$name == 'LAD']

border_gr = import.bed('~/mydata/data/tracks/hg19/cl20161019_LAD_borders_K562.bed')



COL_lad<-c("#FF0000", "#0077FF")
names(COL_lad)<-c('LAD', 'iLAD')
all_data = data.frame()
all_gr = GRanges()
for (prom in names(class_vec)){
    for (bio in unique(map_info$bio_rep)){
        cat(paste0(prom, bio, ':\n'))
        name = map_info[map_info$promoter==prom & map_info$bio==bio, 'name']
        rev_data = rev_table_list[[name]]
        map_gr = get_data_table(rev_data, lad_gr, lmnb1_10kb)
        dist = distanceToNearest(map_gr, border_gr)
        map_gr$border_dist = NaN
        map_gr[queryHits(dist)]$border_dist = mcols(dist)$distance
        map_gr$unique = map_gr$ratio1 > 0.7 & map_gr$ratio2 < 0.1 & map_gr$avg_mapq > 10
        print('LAD/iLAD classification:')
        print(table(map_gr$LAD))
        print('Uniquely mapped:')
        print(table(map_gr$unique))
        print('LAD/iLAD uniquely mapped:')
        print(table(map_gr$LAD[map_gr$unique]))

        name_vec = info[info$promoter==prom & info$bio==bio, 'name']
        rep1 = table_list[[name_vec[1]]]
        rep2 = table_list[[name_vec[2]]]
        bc_vec = names(map_gr)
        count_data = data.frame(row.names=bc_vec,
                                cDNA_cpm=rowMeans(cbind(rep1[bc_vec, 'cDNA_cpm'],
                                                        rep2[bc_vec, 'cDNA_cpm'])),
                                gDNA_cpm=rowMeans(cbind(rep1[bc_vec, 'gDNA_cpm'],
                                                        rep2[bc_vec, 'gDNA_cpm'])))
        gDNA_count = cbind(rep1[bc_vec, 'gDNA_count'], rep2[bc_vec, 'gDNA_count'])
        count_data$min_gDNA = apply(gDNA_count, 1, min)
        size = vector(length=nrow(count_data))
        size[which(count_data$min_gDNA <= 100)] = "gDNA <= 100"
        size[which(count_data$min_gDNA > 100 &
                   count_data$min_gDNA <= 1000)] = "100 < gDNA <= 1000"
        size[which(count_data$min_gDNA > 1000 &
                   count_data$min_gDNA <= 10000)] = "1000 < gDNA <= 10000"
        size[which(count_data$min_gDNA > 10000)] = "gDNA > 10000"
        size_vec = c("gDNA <= 100"=0.1, "100 < gDNA <= 1000"=0.5,
                     "1000 < gDNA <= 10000"=1, "gDNA > 10000"=2)
        count_data$size = factor(size, levels=names(size_vec))
        count_data$norm = count_data$cDNA_cpm / count_data$gDNA_cpm
        count_data$lad = ifelse(map_gr$LAD==1, 'LAD', 'iLAD')
        count_data$lamin = map_gr$lamin
        count_data$border_dist = map_gr$border_dist
        count_data$unique = map_gr$unique
        cd = count_data[count_data$unique & !is.na(count_data$norm) &
                        !is.na(count_data$lamin), ]
        cd$log2 = log2(cd$norm + 0.001)
        lm_fit = lm(cd$log2 ~ -1 + cd$lamin)
        title = paste0(name, '; ', class_vec[prom], '\nlog(expr) = ',
                       round(lm_fit$coefficients[1], 2), ' * log(lamin)')
        cd$prom = prom
        cd$bio = bio
        all_data = rbind(all_data, cd)
        all_gr = c(all_gr, map_gr)
        print(ggplot(cd, aes(x=lad, y=log2, fill=lad)) +
                    ggtitle(paste0(name, '; ', class_vec[prom])) +
                    geom_violin() +
                    scale_fill_manual(values=COL_lad) +
                    ylim(-10, 6.2) +
                    geom_point(aes(size=size), position=position_jitter(width=0.3),
                               alpha=0.5) +
                    scale_size_manual(values=size_vec))
        print(ggplot(cd, aes(x=lamin, y=log2)) +
                    ggtitle(title) +
                    ylim(-10, 6.2) +
                    geom_smooth(method='lm') +
                    geom_point(aes(size=size), alpha=0.8) +
                    scale_size_manual(values=size_vec))
        print(ggplot(cd[cd$lad=='LAD', ], aes(x=log10(border_dist), y=log2)) +
                    ggtitle(title) +
                    geom_point(aes(size=size), alpha=0.8) +
                    scale_size_manual(values=size_vec))
    }
}

all_lad_rvse = all_data[all_data$lad=='LAD'& all_data$prom!='hPGK', ]

all_gr$promoter = all_data[names(all_gr), 'prom']
all_gr$expression = all_data[names(all_gr), 'log2']
all_gr$min_gDNA = all_data[names(all_gr), 'min_gDNA']
all_gr$min_gDNA = all_data[names(all_gr), 'min_gDNA']

export.gff3(all_gr[names(all_gr)%in%rownames(all_lad_rvse)],
            'all_lad_rvse_intergrations.gff3')

high_lad_rvse = all_lad_rvse[all_lad_rvse$log2 > 0, ]
export.gff3(all_gr[names(all_gr)%in%rownames(high_lad_rvse)],
            'high_lad_rvse_intergrations.gff3')

lad_levels = c('iLAD', 'LAD')
pdf('cl20171120_combined_rvse_lad_expresion.pdf', width=5, height=5)
for(prom in names(class_vec)){
    data = all_data[all_data$prom==prom, ]
    data$lad = factor(data$lad, levels=lad_levels)
    lad_names = paste0(lad_levels, '; n=', table(data$lad))
    names(lad_names) = lad_levels
    data$lad_n = lad_names[data$lad]
    COL_lad_n = COL_lad[names(lad_names)]
    names(COL_lad_n) = lad_names
    data_summary = aggregate(log2 ~ lad_n, median, data=data)
    if (prom == 'hPGK') {
        name = 'human PGK promoter'
    } else {
        name = paste0('promoter nr.', prom, '; ', class_vec[prom])
    }
    data$lad = factor(data$lad, levels(c('iLAD', 'LAD')))
    print(ggplot(data, aes(x=lad_n, y=log2, fill=lad_n, ymin=log2, ymax=log2)) +
            geom_violin() +
            ggtitle(name) +
            scale_fill_manual(values=COL_lad_n) +
            ylim(-10, 6.2) +
            geom_point(position=position_jitter(width=0.3),
                       alpha=0.5, size=0.5) +
            theme_bw() +
            xlab('LAD HMM call') +
            ylab('normalized cDNA/gDNA (log2)') +
            geom_crossbar(data=data_summary, color = "green", width=0.3))
}
dev.off()

```
