# knitr document van Steensel lab

# G9a/LBR knockout/knockdown expression
## Christ Leemans, 03-11-2016 - to date

##Introduction
With the help of Joris's SuRE data, Bas defined, by comparing to groCap data, 3 groups of genes in LADs: not repressed, repressed and inactive. After the K562 TTRIP experiments we thought that the position in the LAD might be affecting this behavior.


## LAD position
first let's get some info on the positioning in the LADs.

```r

##tss is just really convenient point, especially with small LAD-states and big genes
load("../raw_data/gencode.sure.160617.rda")
load('../results/cl20161025_G9a_LBR_KD_KO_expression.rData')
gencode.sure.160617$tss_1 = gencode.sure.160617$tss + 1
write.table(gencode.sure.160617[,c('chr', 'tss', 'tss_1', 'name')],
            file='../raw_data/sure_genes.bed', col.names=F, row.names=F, quote=F, sep='\t')

gff_genes$start_site = start(gff_genes)
gff_genes[strand(gff_genes)=='-']$start_site = end(gff_genes[strand(gff_genes)=='-'])
gff_genes$start_site_1 = gff_genes$start_site + 1
write.table(cbind.data.frame(paste0('chr', seqnames(gff_genes)), gff_genes$start_site,
                             gff_genes$start_site_1, gff_genes$gene_id),
            file='../raw_data/knockout_genes.bed', col.names=F, row.names=F, quote=F, sep='\t')
```

```
bedtools sort -i raw_data/sure_genes.bed > raw_data/sure_genes_sorted.bed 
bedtools closest -t all -d -a raw_data/sure_genes_sorted.bed -b /home/NFS/users/c.leemans/data/tracks/hg19/cl20161019_LAD_borders_K562.bed | awk '{print $4"\t"$8}' > raw_data/sure_lad_distances.txt
bedtools intersect -wb -a raw_data/sure_genes_sorted.bed -b /home/NFS/users/c.leemans/data/tracks/hg19/cl20161019_LAD_continuous_2state_K562.bed | \
    awk '{
        print $4"\t"$8"\t"$7-$6

    }' > raw_data/sure_lad_size.txt

bedtools sort -i raw_data/knockout_genes.bed > raw_data/knockout_genes_sorted.bed 
bedtools closest -t all -d -a raw_data/knockout_genes_sorted.bed -b /home/NFS/users/c.leemans/data/tracks/hg19/cl20161019_LAD_borders_K562.bed | awk '{print $4"\t"$8}' > raw_data/knockout_lad_distances.txt
bedtools intersect -wb -a raw_data/knockout_genes_sorted.bed -b /home/NFS/users/c.leemans/data/tracks/hg19/cl20161019_LAD_continuous_2state_K562.bed | \
    awk '{
        print $4"\t"$8"\t"$7-$6

    }' > raw_data/knockout_lad_size.txt
```


```{r}
library(ggplot2)

load("../raw_data/gencode.sure.160617.rda")
load('../results/cl20161025_G9a_LBR_KD_KO_expression.rData')

sure_lad_distance = read.table('../raw_data/sure_lad_distances.txt')
ko_lad_distance = read.table('../raw_data/knockout_lad_distances.txt')

P_distance_match = match(P$name, sure_lad_distance[,1])

P$lad_distance = sure_lad_distance[P_distance_match,2]
ggplot(P, aes(x=class, y=log2(lad_distance), colour=class)) + geom_violin(alpha=0.4) + geom_point(position=position_jitter(width=0.8), alpha=0.3, size=0.3)

ggplot(P, aes(x=log2(lad_distance), colour=class)) + geom_density(adjust=1/3)


gff_distance_match = match(gff_genes$gene_id, ko_lad_distance[,1])

gff_genes$lad_distance = ko_lad_distance[gff_distance_match,2]
ggplot(cbind.data.frame(lad_distance=gff_genes$lad_distance, padj=de_wt_G9a$padj), aes(x=log2(lad_distance), colour=padj<0.05)) + geom_density(adjust=1/3)


```

**conclusion:**

There seems to be a slight tendency for non represed promotors to be more towards the borders. 

```{r}

sure_lad_size = read.table('../raw_data/sure_lad_size.txt')
ko_lad_size = read.table('../raw_data/knockout_lad_size.txt')

P_size_match = match(P$name, sure_lad_size[,1])

P$lad_size = sure_lad_size[P_size_match,3]
ggplot(P, aes(x=class, y=log2(lad_size), colour=class)) + geom_violin(alpha=0.4) + geom_point(position=position_jitter(width=0.8), alpha=0.3, size=0.3)

ggplot(P, aes(x=log2(lad_size), colour=class)) + geom_density(adjust=1/3)


gff_size_match = match(gff_genes$gene_id, ko_lad_size[,1])

gff_genes$lad_size = ko_lad_size[gff_size_match,3]
ggplot(cbind.data.frame(lad_size=gff_genes$lad_size, padj=de_wt_G9a$padj, lad=gff_genes$lad_kbm7), aes(x=padj<0.05, y=log2(lad_size), colour=padj<0.05)) + geom_violin(alpha=0.4) + geom_point(position=position_jitter(width=0.8), alpha=0.3, size=0.3) + facet_wrap(~lad)

ggplot(cbind.data.frame(lad_size=gff_genes$lad_size, padj=de_wt_G9a$padj), aes(x=log2(lad_size), colour=padj<0.05)) + geom_density(adjust=1/3)

```

**conclusion:**

There seems to be a slight tendency for non represed promotors to be more in smaller LADs. This seems to be in concert with previous findings that the size of the LAD adds to the repressive power of the LAD.


```{r}


P$rel_lad_pos = NA
P$rel_lad_pos[!is.na(P$lad_size)] = P$lad_distance[!is.na(P$lad_size)] / P$lad_size[!is.na(P$lad_size)]

ggplot(P, aes(x=class, y=rel_lad_pos, colour=class)) + geom_violin(alpha=0.4) + geom_point(position=position_jitter(width=0.8), alpha=0.3, size=0.3)

ggplot(P, aes(x=rel_lad_pos, colour=class)) + geom_density(adjust=1/3)



gff_genes$rel_lad_pos = gff_genes$lad_distance / gff_genes$lad_size
ggplot(cbind.data.frame(rel_lad_pos=gff_genes$rel_lad_pos, padj=de_wt_G9a$padj, lad=gff_genes$lad_kbm7), aes(x=padj<0.05, y=rel_lad_pos, colour=padj<0.05)) + geom_violin(alpha=0.4) + geom_point(position=position_jitter(width=0.8), alpha=0.3, size=0.3) + facet_wrap(lad)

ggplot(cbind.data.frame(rel_lad_pos=gff_genes$rel_lad_pos, padj=de_wt_G9a$padj), aes(x=rel_lad_pos, colour=padj<0.05)) + geom_density(adjust=1/3)

```

**conclusion:**

This was a bit fruitless....