
# knitr document van Steensel lab

# Main Title
## Christ Leemans, from date - to date 

### Introduction
Introduction to the project

### Description of Data

* Where did I get the data, how does the data look, how many samples and replicates



### Path, Libraries, Parameters and Useful Functions
```{r message=FALSE}

StartTime <-Sys.time()

# 6-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),3,8) 

# libraries:


```

### Custom functions
Functions used


### Load data
Data loading 


```{r}

# Data loading

```


### Some data pre-processing
Calculate some basic values/quality control

```{r}
# pre-processing

```


### Step 1
How reprodicible are the Hi-C data (only dynamic LAD bins):

```{r}
# perform analysis
```
**Conclusion:** the conclusion



### Step 2
description

```{r warning=FALSE, fig.width=8, fig.height=5, out.width=700}
# next step
```
**Conclusion:**
conclusion of step 2


### Session Info
```{r}

sessionInfo()
getwd()
date()
paste("Run time: ",format(Sys.time()-StartTime))

```

</body>
