# knitr document van Steensel lab

# Gene repression in LADs
## Christ Leemans, 03-11-2016 - to date

## Introduction
Generally speaking, genes inside lamina associated domains are not or very lowly expressed. These genes can either be actively repressed by their DNA context (e.g. heterochromatin, lamina association), or simply be inactive (because essential factors for expression are missing?). Yet another group of genes seem to evade gene repression in the context of lamina associated domains. Previously we devided promoters in these three groups. Now I would like to see if there is any particular organisation of these promoters within LADs, are these groups clustering together, or are they interspersed in a way such that the inactive promoters can help repress the repressed promoters.

## Description of Data.

gencode.sure.160617.rda: 
    file from Joris, received 17 June 2016. Promoter positions in this file are from Gencode. Contains SuRE, gro-cap and cage expression data as well as the number of tissues in which each promoter is expressed.


## libraries, paths and data prep


```{r, fig.width=10, fig.height=10}
library(reshape2)
library(rtracklayer)
library(ggplot2)
library(DESeq2)
library(gridExtra)
library(plyr)
library(scales)
library(Matrix)
library(grid)
library(glmnet)
library(rtfbs)
library(BSgenome)
library(BSgenome.Hsapiens.UCSC.hg19)
library(doMC)
library(ggdendro)
registerDoMC(cores=10)
load('../raw_data/biomart.rdata')

## There was one promoter that was wrongly annotated
bm_p[bm_p$ensembl_transcript_id=='ENST00000357491','ensembl_gene_id' ] = 'ENSG00000196350'
```



### SuRE data
Previously, Bas defined three different groups of promoters in LADs. The "escaper", "repressed" and "inactive" group. The "escaper" group contains promoters that are expressed in their endogenous LAD environment. The "repressed" group is inactive in the endogenous setting, but they do show autonomous promoter activity in SuRE data. The 'inactive' group shows neither endogenous expression nor autonomous promoter activity in SuRE data.

```{r, fig.width=10, fig.height=8}
load("../raw_data/gencode.sure.160617.rda")
Prom<-gencode.sure.160617; rm(gencode.sure.160617) #simpler name
#first re-calculate pseudocounts without jitter
P<-Prom[,c(1:8,23,20,26,28, 27)] #SuRE, GRO-cap, CAGE and LAD columns only
names(P)[9:13]<-c("SuRE", "GROcap", "CAGE", "LAD", 'tissues_expressed')


## for promoters and gene expression let's convert promoter transcript id's to gene id's
P$ensembl_transcript_id = do.call(rbind, strsplit(P$name, split='[.]'))[,1]

nrow(P) #orriginal number of rows
bm_match = match(P$ensembl_transcript_id, bm_p$ensembl_transcript_id)
P<-merge(P, bm_p, by="ensembl_transcript_id", all.x=TRUE)
nrow(P) #some double rows were introduced

P = P[match(Prom$name, P$name), ]

length(unique(P$ensembl_gene_id)) #number of unique genes

table(P[,c('strand.x','strand.y')]) #almost all strand listings are consistent

P<-P[, colnames(P)!='strand.y']
colnames(P)[colnames(P)=='strand.x'] = "strand"



## to be used by CGtools as the complete set of TSS's
peaks = data.frame(seqname=P$chr,
                   start=P$tss,
                   end=P$tss,
                   strand=P$strand)


Pseud<-min(P$SuRE[P$SuRE>0], na.rm=TRUE)/2
P$SuRE<-P$SuRE+Pseud
P$SuRE<-log10(P$SuRE)
PseudGro<-min(P$GROcap[P$GROcap>0], na.rm=TRUE)/2
P$GROcap<-P$GROcap+PseudGro
P$GROcap<-log10(P$GROcap)
PseudCage<-min(P$CAGE[P$CAGE>0], na.rm=TRUE)/2
P$CAGE<-P$CAGE+PseudCage
P$CAGE<-log10(P$CAGE)

#then calculate running mean for iLAD promoters:
P<-P[order(P$SuRE,sample(c(1:nrow(P)))),] #sort by SuRE and then random for ties
n<-60 #number of windows
w<-501 #window width (number of datapoints); if n*w > nrow(P) then windows overlap
s<-round(seq(from=w/2+0.0001, to=nrow(P)-w/2, length.out=n))
RM<-data.frame(SuRE.low=rep(NA,n), SuRE.mean=rep(NA,n), SuRE.hi=rep(NA,n), GROcap.lad=rep(NA,n), GROcap.ilad=rep(NA,n))
RM$SuRE.low=P$SuRE[s-floor(w/2)]
for(i in 1:n){RM$SuRE.mean[i]=mean(P$SuRE[(s[i]-floor(w/2)):(s[i]+floor(w/2))], na.rm=TRUE)}
RM$SuRE.hi=P$SuRE[s+floor(w/2)]
for(i in 1:n)
  {t<-P[(s[i]-floor(w/2)):(s[i]+floor(w/2)),]
   RM$GROcap.lad[i]<-mean(t$GROcap[t$LAD==1], na.rm=TRUE)
   RM$GROcap.ilad[i]<-mean(t$GROcap[t$LAD==0], na.rm=TRUE)
  }


#add first datapoint (SuRE equals pseudocount)
RM1<-RM[0,] #empty df
RM1[1,]<-c(rep(log10(Pseud),3), mean(P$GROcap[P$SuRE==log10(Pseud) & P$LAD==1]), mean(P$GROcap[P$SuRE==log10(Pseud) & P$LAD==0]))
RM<-rbind(RM1, RM)
rm(RM1)

#finally calculate LRS for all promoters:
P$LRS<- P$GROcap - approx(x=RM$SuRE.mean, y=RM$GROcap.ilad, xout=P$SuRE, rule=2)$y
#so the more negative the score, the more 'repressed' is the promoter by its chromatin/LAD context


#(arbitrary) cutoffs to define three groups of promoters:
INACT<- P$SuRE< -0.3 & P$LAD==1 & P$GROcap< -2 #inactive
NREP<- P$SuRE> 0 & P$LRS> -0.5 & P$LAD==1 & P$GROcap> -2 #not repressed
REP<- P$SuRE> 0.3 & P$LRS< -1 & P$LAD==1  & P$GROcap< -2 #repressed
Pcnts<-c(length(which(REP)), length(which(NREP)), length(which(INACT)))
names(Pcnts)<-c("repressed", "escaper", "inactive")
BND <- P$LAD==1 & !INACT & !NREP & !REP

#add class annotation column to P:
P$class<-NA
P$class[P$LAD==0]<-"iLAD"
P$class[INACT]<-"inactive"
P$class[NREP]<-"escaper"
P$class[REP]<-"repressed"
P$class[BND] <- "boundary"
P$class = factor(P$class, levels=c('iLAD', 'escaper', 'repressed', 'inactive', 'boundary'))

COLi<-"#00BBFF11" #dot color for iLAD promoters
COL_lad<-c("#FF0000", "#0077FF")
names(COL_lad)<-c('LAD', 'iLAD')

#color vector for plotting:
COL_class<-c("#A020F0", "#FFA500", "#006400", "#7e7e7e", "#0077FF")
names(COL_class)<-c("repressed", "escaper", "inactive", 'boundary', 'iLAD')

COL<-c("#A020F0", "#FFA500", "#006400")
names(COL)<-c("repressed", "escaper", "inactive")

class_names = paste0(levels(P$class), '; n=',table(P$class))
names(class_names) = levels(P$class)
P$class_n_prom = P$class
levels(P$class_n_prom) = class_names
COL_class_n_prom = COL_class[names(class_names)]
names(COL_class_n_prom) = class_names

lad_names = c(LAD=paste0('LAD; n=', table(P$LAD)['1']),
              iLAD=paste0('LAD; n=', table(P$LAD)['0']))
P$lad_n_prom = factor(ifelse(P$LAD==1, lad_names['LAD'], lad_names['iLAD']))
COL_lad_n = COL_lad
names(COL_lad_n) = lad_names


RM_melt = melt(RM, measure.vars=c('GROcap.ilad', 'GROcap.lad'))
RM_melt$variable = ifelse(RM_melt$variable=='GROcap.lad', lad_names['LAD'], lad_names['iLAD'])
ggplot(P, aes(x=SuRE, y=GROcap, color=lad_n_prom)) +
    geom_point(data=P[P$LAD==0, ], size=0.5, alpha=0.05) + 
    geom_point(data=P[P$LAD==1, ], size=0.5, alpha=0.2) + 
    theme_bw() +
    geom_line(data=RM_melt, aes(x=SuRE.mean, y=value, color=variable), size=1) +
    labs(y='log10(GROcap)', x='log10(SuRE)') +
    theme(legend.title=element_blank()) +
    scale_color_manual(values=COL_lad_n)
p_classes = P[which(P$class %in% c('inactive', 'escaper', 'repressed')),]
ggplot(P, aes(x=SuRE, y=GROcap)) + 
    geom_point(size=0.1,color=COLi) + 
    geom_point(data=p_classes, aes(color=class_n_prom), size=0.6) + 
    labs(y='log10(GROcap)', x='log10(SuRE)') +
    theme_bw() +
    geom_line(data=RM, aes(x=SuRE.mean, y=GROcap.ilad), color=COL_lad['iLAD']) +
    theme(legend.title=element_blank()) +
    scale_colour_manual(values=COL_class_n_prom) 

```

**conclusion:**

We now have a definition of 3 different groups. We can now look at what is different between these groups.


## create GRanges

```{r}
#genes as Granges object
gene_gr <-makeGRangesFromDataFrame(data.frame(seqnames=P$chr,
                                              start=P$txStart,
                                              end=P$txEnd,
                                              strand=P$strand,
                                              class=P$class),
                                              keep.extra.columns=TRUE)

lad_gr = import.bed('../../../data/tracks/hg19/cl20161019_LAD_continuous_2state_K562.bed')

lad_class_gr = gene_gr[gene_gr$class %in% names(COL)]

o = findOverlaps(lad_gr, lad_class_gr)